Re-analysing the data from Moffatt et al.~(2020): A textbook illustration of the fallacy of the null hypothesis

### 2020-09-29 reanalysis.Rmd
bf simulation

---
title: "Re-analysing the data from Moffatt et al. (2020): A textbook illustration of the fallacy of the null hypothesis"
shorttitle: "The fallacy of the null"
author: 
  - name: Ladislas Nalborczyk
    affiliation: 1
    corresponding: yes
    address: GIPSA-lab, CNRS, Univ. Grenoble Alpes, 11 Rue des Mathématiques, 38400 Saint-Martin-d'Hères, France
    email: ladislas.nalborczyk@gipsa-lab.fr
affiliation:
  - id: 1
    institution: Univ. Grenoble Alpes, CNRS, Grenoble INP, GIPSA-lab, 38000 Grenoble, France
abstract: "Moffatt et al. (2020) reported the results of an experiment (N = 26 in the final sample) comparing the facial (surface) electromyographic correlates of mental rumination and distraction, following an experimentally induced stressful event. Based on the observation... We show ou this conclusion seems at best hasty... Indeed, concluding on the null, relation between self-reports and physiological measures... Although the question Moffatt et al. (2020) pursued if of interest, and mehods... it is unclear whether this contribution really adds to the current knowledge... Complete source code, reproducible analyses, and figures are available at https://github.com/lnalborczyk/inner_experience_EMG."
keywords: NHST, Bayesian, fallacy, reanalysis, inner speech, rumination, electromyography
csl: csl/apa7.csl
classoption: man
linenumbers: yes
floatsintext: yes
link-citations: true
numbersections: false
bibliography: [bib/brms_tutorial.bib, bib/r-references.bib]
nocite: | 
  @R-base, @R-knitr, @R-rmarkdown, @R-tidyverse, @R-papaja, @R-wordcountaddin,
  @R-here
output:
  papaja::apa6_pdf:
    keep_tex: true
    highlight: tango
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

>>>>

---
title: "Re-analysing the data from Moffatt et al. (2020): A textbook illustration of the fallacy of the null hypothesis"

shorttitle: "The fallacy of the null"

author: 
  - name: Ladislas Nalborczyk
    affiliation: 1
    corresponding: yes
    address: GIPSA-lab, CNRS, Univ. Grenoble Alpes, 11 Rue des Mathématiques, 38400 Saint-Martin-d'Hères, France
    email: ladislas.nalborczyk@gipsa-lab.fr
affiliation:
  - id: 1
    institution: Univ. Grenoble Alpes, CNRS, Grenoble INP, GIPSA-lab, 38000 Grenoble, France

abstract: "Moffatt et al. (2020) reported the results of an experiment (N = 26 in the final sample) comparing the facial (surface) electromyographic correlates of mental rumination and distraction, following an experimentally induced stressor. Based on an absence of significant difference in the perioral EMG-amplitude between the rumination and distraction conditions, Moffatt et al. (2020) concluded that (the *self-reported*) inner experience was unrelated to peripheral muscular activity as assessed using surface electromyography. We suggest this conclusion is hasty and based on waggly evidence. Indeed, concluding on the absence of an effect based on a low-powered statistical test is strongly problematic/uninformative. Moreover, there relation between self-reports and physiological measures... Although the question Moffatt et al. (2020) pursued is of interest, and mehods... Given these limitations, it is unclear whether the target article really adds to the current knowledge. Complete source code, reproducible analyses, and figures are available at https://github.com/lnalborczyk/inner_experience_EMG."

keywords: NHST, Bayesian, fallacy, reanalysis, inner speech, rumination, electromyography

csl: csl/apa7.csl
classoption: man
linenumbers: yes
floatsintext: yes
link-citations: true
numbersections: false
bibliography: [bib/brms_tutorial.bib, bib/r-references.bib]

nocite: | 
  @R-base, @R-knitr, @R-rmarkdown, @R-tidyverse, @R-papaja, @R-wordcountaddin,
  @R-here

output:
  papaja::apa6_pdf:
    keep_tex: true
    highlight: tango
    latex_engine: xelatex
---

### 2020-09-29 reanalysis.tex
bf simulation

\abstract{
Moffatt et al.~(2020) reported the results of an experiment (N = 26 in the final sample) comparing the facial (surface) electromyographic correlates of induced rumination and distraction, following a stressful event. Based on the observation\ldots{} We show ou this conclusion seems at best hasty\ldots{} Indeed, concluding on the null, relation between self-reports and physiological measures\ldots{} Although the question Moffatt et al.~(2020) pursued if of interest, and mehods\ldots{} it is unclear whether this contribution really adds to the current knowledge\ldots{} Complete source code, reproducible analyses, and figures are available at \url{https://github.com/lnalborczyk/inner_experience_EMG}.
}

@@

Wordcount (excluding abstract, references, tables, and figures): 107

@@

>>>>

\abstract{
Moffatt et al.~(2020) reported the results of an experiment (N = 26 in the final sample) comparing the facial (surface) electromyographic correlates of mental rumination and distraction, following an experimentally induced stressor. Based on an absence of significant difference in the perioral EMG-amplitude between the rumination and distraction conditions, Moffatt et al.~(2020) concluded that (the \emph{self-reported}) inner experience was unrelated to peripheral muscular activity as assessed using surface electromyography. We suggest this conclusion is hasty and based on waggly evidence. Indeed, concluding on the absence of an effect based on a low-powered statistical test is strongly problematic/uninformative. Moreover, there relation between self-reports and physiological measures\ldots{} Although the question Moffatt et al.~(2020) pursued is of interest, and mehods\ldots{} Given these limitations, it is unclear whether the target article really adds to the current knowledge. Complete source code, reproducible analyses, and figures are available at \url{https://github.com/lnalborczyk/inner_experience_EMG}.
}

@@

Wordcount (excluding abstract, references, tables, and figures): 184

@@

\begin{quote}
``In order to test this, a Bayesian paired samples t-test was conducted for the peak log values of muscle activity between the rumination and distraction conditions. This revealed strong evidence in favour of the alternative hypothesis for the FRO muscle (B10 = 18.79), and moderate evidence in favour of the null hypothesis for the OOS (B10 = 0.232) and OOI (B10 = 0.278) muscles, according to current guidelines for interpreting Bayes factors {[}43{]}.''
\end{quote}
